---
title: "Aids Drug"
author: "Rebecca Castro"
date: "9/28/2022"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
triUmp <-read.csv(file="triUmpsal.csv",header=TRUE)

Females=filter(triUmp,gender== "F")
DrugUsers=filter(triUmp,DrugUser=="yes")
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative aids drug (triUmpsal) is under test in clinical trials. Consider the DrugUsers, does this drug seem effective on this group?  Hint - use dplyr to obtain that sub-population with the filter command.  Next, repeat the analysis on just females, does this drug seem effective on women?

<!-- Be sure to produce Descriptive Results and Inferential Results for both groups. !-->

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed, for example sub-setting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

The questions are 1. is the vaccine effective for drug users. 2. Is the vaccine effective for women.

$H_0$:The vaccine is not effective for drug users
$H_A$:The vaccine is effective for drug users

$H_0$: The vaccine is not effective for women
$H_A$: The vaccine is effective for women

# Results

<!--Divide this section into two sub-sections:  One for your descriptive results and one for your inferential results.!-->

## Descriptive Results for druguser

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe patterns, if any, that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming to or departing from what is expected by the null hypothesis !-->

```{r}
barchartGC(~infected + treatment,data=DrugUsers)
barchartGC(~infected + treatment,data=DrugUsers, type="percent")

```

Drug users treated with the vaccine are less likely to get AIDS.

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns, if any, that exist in the data as focused toward the research question!-->
```{r}
table1 <- xtabs(~infected + treatment, DrugUsers)
rowPerc(table1)
colPerc(table1)
```

of all the people who were infected drug users showed 25% had the vaccine and 75% had the placebo and this suggest that the vaccine is effective for drug users.

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfies the assumptions and is appropriate to answer the research question!-->
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The p-value is 1.242e-07 which is less than 0.05 we reject the null. the vaccine is effective in the druguser community.
  
## Descriptive Results for women

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe patterns, if any, that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming to or departing from what is expected by the null hypothesis !-->

```{r}
barchartGC(~infected + treatment,data=Females)
barchartGC(~infected + treatment,data=Females, type="percent")

```

Females treated with the vaccine are more likely to get AIDS.

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns, if any, that exist in the data as focused toward the research question!-->
```{r}
table1 <- xtabs(~infected + treatment, Females)
rowPerc(table1)
colPerc(table1)
```

of all the people who were infected females showed 60% had the vaccine and 40% had the placebo and this suggest that the vaccine is not effective for women.

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfies the assumptions and is appropriate to answer the research question!-->
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The p-value is 7.686e-05 which is less than 0.05 we reject the null. the vaccine increases the probability of women getting AIDS  
  
  
# Discussion and Conclusion

The vaccine is effective for drug users and is not effective for women. 
<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variables causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the George Town videos collection.!

Report on both Drug Users and Females.  Be sure to include the odds for both groups.
-->

 
